---
title: "RegressionAnalysis"
author: "Kevin Thompson"
date: "August 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggfortify)
load("data/.RData")
```

A quick check of the imputed Lot Frontage values before we begin.

```{r imputationcheck, echo=FALSE}
housePrices_imputed_new <- housePrices_imputed %>%
                            filter(Neighborhood %in% c("NAmes", "BrkSide", "Edwards"))

housePrices_imputed_new %>%
  ggplot(aes(x=LotFrontage, y=SalePrice, color=any_missing)) +
  geom_point() +
  labs(x = "Lot Frontage", y = "Price", color = "Missingness", title = "Price vs Frontage")

```

We look 





We fit regression models to get the following residual plots:

```{r modelfit, echo=TRUE}
baseModel <- housePrices_imputed_new %>%
              lm(formula = SalePrice ~ GrLivArea + Neighborhood + Neighborhood*GrLivArea)
noOutlierModel <- lm(data = housePrices_imputed_new[-c(339),], formula = SalePrice ~ GrLivArea + Neighborhood + Neighborhood*GrLivArea)
logModel <- lm(data = housePrices_imputed_new, formula = log(SalePrice) ~ GrLivArea + Neighborhood + Neighborhood*GrLivArea)
lognoOutlierModel <- lm(data = housePrices_imputed_new[-c(339),], formula = log(SalePrice) ~ GrLivArea + Neighborhood + Neighborhood*GrLivArea)
```



```{r residplots, echo=FALSE}
autoplot(baseModel, which = 1:6, ncol = 2, label.size = 3, title = "linear")
autoplot(logModel, which = 1:6, ncol = 2, label.size = 3, title = "log-linear")
```

```{r, echo=TRUE}
summary(baseModel)
summary(noOutlierModel)
summary(logModel)
summary(lognoOutlierModel)
```

It appears that the outlier has significant influence on our coefficient estimates for edwards and the interaction between edwards and gross living area. Though the observation is still influential after taking a log transform, we find that the influence is reduced such that we do not get different answers with or without the observation. So our final model is the log-linear model. 